---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```


# The dvir (Disaster Victim Identification) library

We assume DNA profiles are available from victim samples (post mortem, pm data) and 
reference families (ante mortem, am, data) with missing persons (MP-s). There may be 
several samples from the same victim, potentially of low quality leading to *drop-outs*.
The problem is to identify the MP-s. Some (or all) victims may not be 
among the MP-s. Similarly, there may be MP-s not in the list of victims.
A search strategy
is implemented. All victims are initially tried, one at a time, in all MP positions.
Results are sorted according to the likelihood and assignments with a LR (compared to the null
likelihood) below a user specified limit are omitted from further search. If mutations are 
modelled all LR-s will typically be positive and the limit must be specified to a negative number
to include all possibilities in the future search. Based on this initial screening, all possible assignments
of victims are generated. Note that only a subset, possibly none, of victims may be mapped to MP-s.
The resulting list of assignments may be prohibitively large and for this reason it possible to restrict the search by
specifying that only the `nbest` assignments for each victim be considered.

## Installation

To get the latest version, install from GitHub as follows:

```{r, eval = FALSE}
 # First install devtools if needed
if(!require(devtools)) install.packages("devtools")

# Install dvir from GitHub
devtools::install_github("thoree/dvir")
```
The implementation relies heavily on 
the `ped suite` of R-libraries, in particular `forrel` and `pedmut`. These are automatically installed by the above command.

## Load libraries

```{r}
library(dvir)
```

## Example 1

We consider the following example

```{r, fig.height = 3, fig.width = 7, echo = T}
library(pedtools)
n = 7
ids.from = paste("V", 1:n, sep = "")
sex = c(rep(1, n-1), 2)
df = data.frame(id = ids.from, fid = 0, mid = 0, sex = sex,
                a1 = c(1,2,1,1,2,2,2),
                a2 = c(1,2,1,1,2,2,2))
locus_annotations = list(alleles = 1:3, afreq = c(1, 1, 1)/3)
from = as.ped(df, locusAttributes = locus_annotations)
to = nuclearPed(3, father = "R1", mother = "R2", children= c("MP1","MP2","MP3"))
m = marker(to, "R1" = 1, "R2" = 1,   alleles = 1:3, afreq = c(1, 1, 1)/3, name = "a1")
to = addMarkers(to, m)
plotPedList(list(from, to), marker = 1, 
            titles = c("PM data. 7 victims", "AM data. 3 MP-s"))
```

We do not consider mutations or other artefacts. We assume that copies of victim samples have been identified and merged so that without extra information like (*todo:* add age),
there are six symmetric solutions. If we knew the age ordering,
for instance $age(V1) > age(V2) > age(V3)$, then 
$MP1=V1, MP2=V2, MP3=V3$. Recall, the convention that individuals are
ordered left to right in the pedigree based on age.

### The number of assignments

The number of assignments is

```{r}
m = ncomb(1,0,6,3)
m
```

as is checked by

```{r} 
MPs = c("MP1", "MP2","MP3")
# see function generateMoves below
moves = list(V1 = c("V1", MPs ), V2 = c("V2", MPs ), V3 = c("V3", MPs ), V4 = c("V4", MPs),
             V5 = c("V5", MPs ), V6 = c("V6", MPs ), V7 = "V7" )
a = expand.grid.nodup(moves)
stopifnot(length(a) == m)
```

### The search

The search and the ten best solutions among all possible
```{r}
res = global(from, to, MPs, moves = NULL, limit = -1, verbose = F)
res[1:10,]
```


We can start by limiting the search. First all sex consistent 
marginal moves are generated. 

```{r}
moves = generateMoves(from, to, MPs)
```


We only keep two best marginal candidates

```{r}
moves2 = marginal(from, to,  MPs, moves, limit = -1, sorter = T,  nkeep = 3)
res = global(from, to, MPs, moves = moves2[[1]], limit = -1, verbose = F)
res[1:6,]
```

## Example 2

We next consider a larger example data. First the data is loaded:

```{r}
con <- url("http://familias.name/BookKETP/Files/Grave.RData") 
load(con) 
close(con) # Finished loading data: from, to, ids.to and moves
rm(con)
```

The family with the missing persons MP1-MP8 and
references R1-R5, the genotyped family members, is shown below

```{r, echo = T}
labnew = labels(to)
labnew[c(1:5,7:8,16,20:21)] = " "
refs = paste("R", 1:5, sep = "")
mps = paste("MP", 1:8, sep ="")
plot(to,labs = labnew,  aff = c(refs, mps),
     col = list( red = refs, blue = mps), deceased = mps,
     title = "AM data")
```

The variable `from` is a list of singletons with female victims V1, V3, V4, V5, V6
and male victims V2, V7, V8. The apriori possible number of assignments
ignoring symmetries is

```{r}
ncomb(5, 5, 3, 3)
```
We restrict the number of assignments by requiring $LR > 0.99$ for *marginal* moves.
For instance, based on the below, the possibility `V1 = MP1` will be considered
since the $LR$ comparing the assignment `V1 = MP1, no more victims identified` to the null hypotheses,
`none identified`, exceeds 0.99. 

```{r}
moves = generateMoves(from, to, ids.to)
m = marginal(from , to, ids.to, limit = 0.99, moves = moves, nkeep = 2, sorter = T)
m[[1]]
```
Based on the corresponding LRs below, we realise that `limit = 0` would be a better option, since
the only additions are `V8 = MP8`and `V2 = MP8`

```{r}
m[[2]]
```

However, changing to `limit = 0` considerably increases computation time, 
I haven't checked how much,  but I rather use the list obtained above:

```{r}
res1 = global(from, to, ids.to, limit = 0.99, moves = m [[1]])
head(res1)
```

We check the assignment with
the identification MP8 = V8 added

```{r}
res2 = global(from, to, ids.to, 
       moves = list(V1 = "MP1", V2 = "MP2", V3 = "MP3", V4 = "MP4",
                    V5 = "MP5", V6 = "MP6", V7 = "MP7", V8 = "MP8"))
res2
exp(res2$loglik-res1$loglik[1])
```

Finally, we would like to do an exhaustive search and the code is below. Unfortunately,
we may have to wait forever for the solution with the current implementation

```{r, cache =T, eval  = F}
res3 = global(from, to, ids.to, moves = NULL)
head(res3)
```



