---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.align = "center",
  dpi = 300,
  message = FALSE, warning = FALSE
)
```

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/dvir)](https://CRAN.R-project.org/package=dvir)
[![](https://cranlogs.r-pkg.org/badges/grand-total/dvir?color=yellow)](https://cran.r-project.org/package=dvir)
[![](https://cranlogs.r-pkg.org/badges/last-month/dvir?color=yellow)](https://cran.r-project.org/package=dvir)
<!-- badges: end -->

# Disaster Victim Identification in R

The purpose of **dvir** is to implement state-of-the-art algorithms for DNA-based disaster victim identification (DVI). In particular, **dvir** performs *joint* identification of multiple victims.

The methodology and algorithms of **dvir** are described in [Vigeland & Egeland (2021): DNA-based Disaster Victim Identification](https://www.nature.com/articles/s41598-021-93071-5).

The **dvir** package is part of the [pedsuite](https://magnusdv.github.io/pedsuite/), a collection of R packages for pedigree analysis. Much of the machinery behind **dvir** is imported from other pedsuite packages, especially **pedtools** for handling pedigree data, and **forrel** for the calculation of likelihood ratios. A comprehensive presentation of these packages, and much more, can be found in the book [Pedigree Analysis in R](https://shop.elsevier.com/books/pedigree-analysis-in-r/vigeland/978-0-12-824430-2).

## Installation

To get the current official version of **dvir**, install from CRAN as follows:
```{r, eval = FALSE}
install.packages("dvir")
```

Alternatively, the latest development version may be obtained from GitHub:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("magnusdv/dvir")
```

```{r, echo = F}
library(dvir)
```

## Example
We illustrate joint DVI analysis using a toy example from the paper (Vigeland & Egeland, 2021), shown in the plot below. Three victim samples (V1, V2, V3) are to be matched against three missing persons (M1, M2, M3) in two different families.

```{r example-plot1, echo = FALSE, fig.height = 3, fig.width = 6.5, out.width = "85%"}
plotDVI(example2, marker = 1, titles = c("PM data", "AM data"), frames = FALSE)
```

The hatched symbols indicate genotyped individuals. In this simple example we consider only a single marker, with 10 equifrequent alleles denoted 1, 2,..., 10. The available genotypes are shown in the figure. 

DNA profiles from victims are referred to as _post mortem_ (PM) data, while the _ante mortem_ (AM) data contains profiles from the reference individuals R1 and R2.

### Assignments
A possible solution to the DVI problem is called an _assignment_. In our toy example, there are _a priori_ 14 possible assignments respecting the individuals' sex:

```{r, echo = F}
generatePairings(example2) |> expand.grid.nodup() |>
  knitr::kable(align = "c")
```

Each row indicates the missing persons corresponding to V1, V2 and V3 (in that order) with `*` meaning _not identified_. For example, the first line contains the _null model_ corresponding to none of the victims being identified, while the last line gives the assignment where `(V1, V2, V3) = (M1, M2, M3)`, 
For each assignment `a` we can calculate the likelihood, denoted `L(a)`. The null likelihood is denoted `L0`.

Our aim in this brief tutorial is to perform the following tasks using **dvir**:

1) Rank the assignments according to how likely they are. 
    For each assignment `a` we compute the LR comparing it to the null model: `LR = L(a)/L0`. 
2) Find the *posterior pairing probabilities* `P(V_i = M_j | data)` for all pairs of victims and missing persons.

### The data

The pedigrees and genotypes for this toy example are available within **dvir** as a built-in dataset, under the name `example2`.

```{r, eval = F}
library(dvir)
```

```{r}
example2
```

Internally, all DVI datasets in **dvir** have the structure of a list, with elements `pm` (the victim data), `am` (the reference data) and `missing` (a vector naming the missing persons):
We can inspect the data by printing each object. For instance, in this case `am` is a list of two pedigrees:
```{r}
example2$am
```
Each pedigree is printed in so-called _ped format_, with columns `id` (ID label), `fid` (father's ID, or `*` if no father is included), `mid` (mother's ID), `sex` (1 = male; 2 = female) and `L1` (genotypes at locus `L1`).

The code generating this dataset can be found in the github repository of **dvir**, more specifically here: https://github.com/magnusdv/dvir/blob/master/data-raw/example2.R.

A great way to inspect a DVI dataset is to plot it with the function `plotDVI()`.
```{r example-plot2, fig.height = 3, fig.width = 7, out.width = "80%"}
plotDVI(example2)
```

The `plotDVI()` function offers many parameters for tweaking the plot; see the help page `?plotDVI()` for details.

### Joint identification

The `jointDVI()` function performs joint identification of all three victims, given the data. It returns a data frame ranking all assignments with nonzero likelihood:

```{r}
jointRes = jointDVI(example2, verbose = FALSE)

# Print the result
jointRes
```

The output shows that the most likely joint solution is (V1, V2, V3) = (M1, M2, M3), with an LR of 250 compared
to the null model. 

The function `plotSolution()` shows the most likely solution:
```{r solution, fig.height = 3.5, fig.width = 7, out.width = "75%"}
plotSolution(example2, jointRes, marker = 1)
```

By default, the plot displays the assignment in the first row of `jointRes`. To examine the second most likely, add `k = 2` (and so on to go further down the list).

### Posterior pairing probabilities
Next, we compute the posterior pairing (and non-pairing) probabilities. This is done by feeding the output from `jointDVI()` into the function `Bmarginal()`.

```{r}
Bmarginal(jointRes, example2$missing, prior = NULL)
```
Here we used a default flat prior for simplicity, assigning equal prior probabilities to all assignments.

We see that the posterior pairing probabilities for the most likely solution are

  * _P_(V1 = M1 | data) = 0.88,
  * _P_(V2 = M2 | data) = 0.95,
  * _P_(V3 = M3 | data) = 0.83.

----
