---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```


# The dvir (Disaster Victim Identification) library

DNA profiles are available from victims (post mortem, pm, data) and 
reference families (ante mortem, am, data) with missing persons (mp-s). The problem 
is to identify the mp-s. Some (or all) victims may not be 
among the mp-s. Similarly, there may be mp-s not in the list of victims.
A search strategy
is implemented. All victims are initially tried, one at a time, in all mp positions.
Results are sorted according to the likelihood and moves with a LR (compared to the null
likelihood) below a user specified limit are omitted from further search. If mutations are 
modelled all LR-s will typically be positive and the limit must be specified to a negative number
to include all possibilities in the future search. Based on this initial screening, all possible moves
of victims are generated. Note that only a subset, possibly none, of victims may be mapped to mp-s.
The resulting list of moves may be prohibitively large and for this reason it possible to restrict the search by
specifying that only the `nbest` moves for each victim be considered.

In addition, a forward stepwise approach, conceptually similar to variable selection in regression analysis, is implemented in the function `forward`. This approach is generally faster, but may fail to find a solution.

## Installation

To get the lastest version, install from GitHub as follows:

```{r, eval = FALSE}
 # First install devtools if needed
if(!require(devtools)) install.packages("devtools")

# Install mut2 from GitHub
devtools::install_github("thoree/dvir")
```
The implementation relies heavily on 
the `pedtools` suite of R-libraries, in particular the `forrel` and `pedmut`
libraries
which can be installed by running

```{r, eval = FALSE}
devtools::install_github("magnusdv/forrel")
devtools::install_github("magnusdv/pedmut")
```
                        

## Example 1

The number of combinations

```{r example1, message = FALSE}
library(dvir)
ncomb(3,3,1,1)
```

