---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.align = "center",
  dpi = 300,
  message = FALSE, 
  warning = FALSE
)
```

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/dvir)](https://CRAN.R-project.org/package=dvir)
[![](https://cranlogs.r-pkg.org/badges/grand-total/dvir?color=yellow)](https://cran.r-project.org/package=dvir)
[![](https://cranlogs.r-pkg.org/badges/last-month/dvir?color=yellow)](https://cran.r-project.org/package=dvir)
<!-- badges: end -->


# Disaster Victim Identification in R

The purpose of **dvir** is to implement state-of-the-art algorithms for DNA-based disaster victim identification (DVI). In particular, **dvir** performs *joint* identification of multiple victims.

The methodology and algorithms of **dvir** are described in

* Egeland & Vigeland (2025): [Kinship cases with partially specified hypotheses](https://doi.org/10.1016/j.fsigen.2025.103270)
* Vigeland & Egeland (2021): [Joint DNA-based Disaster Victim Identification](https://www.nature.com/articles/s41598-021-93071-5)

The **dvir** package is part of the [pedsuite](https://magnusdv.github.io/pedsuite/), a collection of R packages for pedigree analysis. Much of the machinery behind **dvir** is imported from other pedsuite packages, especially **pedtools** for handling pedigree data, and **forrel** for the calculation of likelihood ratios. A comprehensive presentation of these packages, and much more, can be found in the book [Pedigree Analysis in R](https://shop.elsevier.com/books/pedigree-analysis-in-r/vigeland/978-0-12-824430-2).

## Installation

To get the current official version of **dvir**, install from CRAN as follows:

```{r, eval = FALSE}
install.packages("dvir")
```

Alternatively, the latest development version may be obtained from GitHub:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("magnusdv/dvir")
```

## Joint DVI analysis

```{r}
library(dvir)
```

We illustrate joint DVI analysis using a toy example from Vigeland & Egeland (2021), shown below. Three victim samples (V1, V2, V3) are to be matched against three missing persons (M1, M2, M3) from two families.

```{r example2-plot, echo = FALSE, fig.height = 3, fig.width = 6.5, out.width = "85%"}
plotDVI(example2, marker = 1, frames = FALSE, style = 2)
```

The hatched symbols indicate genotyped individuals. This simple example uses a single marker with 10 equifrequent alleles denoted 1, 2, ..., 10. The observed genotypes are shown in the figure.

DNA profiles from victims are referred to as _post mortem_ (PM) data, while the _ante mortem_ (AM) data contain profiles from the reference individuals, R1 and R2.

The function `dviJoint()` performs a joint likelihood search over all valid assignments.

```{r example2-joint}
joint = dviJoint(example2)
```

The output contains the log-likelihood of each of the 14 assignments, and likelihood ratios (LRs) comparing the most likely solution with each of the others. Thus the first row has `LR = 1`, while larger values indicate less likely assignments:

```{r, echo = F}
knitr::kable(joint, row.names = T, digits = 3)
```

The columns V1, V2 and V3 give the PM-oriented assignment: each victim is either paired with a missing person or remains unassigned (`"*"`). The columns M1, M2 and M3 give the AM-oriented view of the same assignments.

We see that the most likely joint solution is V1 = M1, V2 = M2 and V3 = M3.

## Automated DVI workflow

For complete DVI analysis, the main function is `dviSolve()`. It combines several steps, including exclusions, undisputed identifications, pairwise LR calculations, joint analysis of unresolved families, and conclusions based on generalised likelihood ratios (GLRs).

We illustrate this with the built-in dataset `fire`, which includes three victim samples and three missing persons from the same family. The function `plotDVI()` gives an overview:

```{r fire-plot, fig.height = 3.5, fig.width = 6.5, out.width = "85%"}
plotDVI(fire, style = 2)
```

Here is the complete workflow using LR threshold 10 000:
```{r fire-solve, R.options = list(width = 100)}
res = dviSolve(fire, detailedOutput = TRUE, threshold = 1e4, verbose = TRUE)
```

The output contains AM-oriented and PM-oriented summary tables, including conclusions and comments. With `detailedOutput = TRUE`, the result also contains the pairwise LR matrix, the exclusion matrix, and the joint tables used in the analysis:

#### `res$AM`

```{r, echo = F}
knitr::kable(res$AM, digits = 2)
```

#### `res$PM`

```{r, echo = F}
knitr::kable(res$PM, digits = 2)
```

#### `res$LRmatrix`

```{r, echo = F}
knitr::kable(res$LRmatrix, digits = 3)
```

#### `res$exclusionMatrix`

```{r, echo = F}
knitr::kable(res$exclusionMatrix)
```

#### `res$jointTables` (truncated)

```{r, echo = F}
knitr::kable(res$jointTables[[1]] |> head(), digits = 2)
```

### Conclusion

In the `fire` example, V1 is identified as M1 by joint GLR analysis (Egeland & Vigeland, 2025). The two remaining victims are matched to the two missing full siblings M2 and M3, but the data do not distinguish between the two sibling assignments. This is reported as a symmetric match.
