---
title: "Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Some examples are discussed based on the following data from
Kling et al (2021): `KETPch4`, `KETPex481`, `KETPex497`, and `KETPex497`.

```{r setup, echo = F}
library(dvir)
```

## Example 4.8.4, Kling et al. (2021)

This example considers several reference families, see below figure. There are 21 markers and we use an equal mutation model with mutation rate 0.005. We assume equal mutation rates females and males for all markers
in all examples involving mutation. 

```{r fig1, fig.height = 3, fig.width = 7, out.width = "100%", fig.cap = "Example 4.8.4, Figure 4.15, Kling et al. (2021) with genotypes for the two first markers."}
plotDVI(KETPch4, nrowPM = 2,  marker = 1:2)
```
There are 49 possible solutions as we find from

```{r,eval = F}
ncomb(2,2,2,2)
```

The pairwise LR approach, similar to what is implemented in e.g., Familias gives
```{r}
pairwiseLR(KETPch4)$LRmatrix
```
There is some, albeit not strong, indication that (V1 = MP1, V2 = MP2, V3 = MP3, V4 = MP4). However,  a convincing solution emerges from the joint approach
```{r}
res1 = jointDVI(KETPch4, disableMutation = F, verbose = F)
res1[1:3,]
```
The solution can be plotted as follows
```{r, fig.height = 3, fig.width = 7, out.width = "100%", fig.cap = "Figure 2. Solution of joint approach in Example 1."}
plotSolution(KETPch4, res1)
```

We can also compute posterior pairing and non-pairing probabilities
```{r}
Bmarginal(res1, KETPch4$missing)
```


## Example 4.8.1, Kling et al. (2021)

The problem is shown in Figure 3 below:
```{r, fig.height = 3, fig.width = 7, out.width = "100%", fig.cap = "Figure 3. Example 4.8.1, Figure 4.13, Kling et al. (2021)"}
plotDVI(KETPex481, nrowPM = 2, widths = c(1, 4.5), cex = 1.0)
```
In this case there are 21 markers, but mutations are not modelled.
The pairwise LR approach does not give a clear conclusion:
```{r}
pairwiseLR(KETPex481)$LRmatrix
```
However,  a convincing solution emerges from the joint approach
```{r}
res2 = jointDVI(KETPex481, verbose = F)
res2
```
The solution can be plotted as follows
```{r, fig.height = 3, fig.width = 7, out.width = "100%", fig.cap = "Figure 4. Solution of joint approach in Example 2."}
plotSolution(KETPex481, res2)
```

## Exercise 4.9.7, Kling et al. (2021)

The problem is shown in Figure 3 below:
```{r, fig.height = 3, fig.width = 7, out.width = "100%", fig.cap = "Figure 5. Exercise 4.9.7, Figure 4.22, Kling et al. (2021)"}
plotDVI(KETPex497, nrowPM = 3,   cex = 1.1)
```
There are 23 markers, and an equal mutation model with rate 0.001.
Part (a) of the exercise asks for a list of possible assignments. 
There are `ncomb(3, 3, 0, 0)` = `r ncomb(3, 3, 0, 0)` assignments.

The list of assignments (not shown) can be generated by
```{r, eval = F}
lst = generatePairings(KETPex497)
expand.grid.nodup(lst)
```
Then, the joint solution is required
```{r}
res3 = jointDVI(KETPex497, disableMutations = FALSE,  verbose = F)
res3
```

The solution can be plotted as follows
```{r, fig.height = 3, fig.width = 7, out.width = "100%", fig.cap = "Figure 4. Solution of joint approach in Exercise497."}
plotSolution(KETPex497, res3)
```

## Exercise 4.9.8, Kling et al. (2021)

Data used in Exercise 4.9.8 in Kling et al. (2021) . There are 2 female victims and 1 male. There is one reference family with 2 missing females and one missing male. There are 16 markers, equal mutation model, rate 0.001.

```{r fig498, fig.height = 3, fig.width = 7, out.width = "100%", fig.cap = "Example 4.9.8, Kling et al. (2021) with genotypes for the two first markers."}
plotDVI(KETPex498, nrowPM = 3,   cex = 1, widths = c(1,4))
```

The result of a pairwise search is again not conclusive
```{r}
pairwiseLR(KETPex498)$LRmatrix
```
whereas a joint search is convincing
```{r}
res = jointDVI(KETPex498, verbose = F)
res
```
Here's the solution plot
```{r, fig.height = 3, fig.width = 7, out.width = "100%", fig.cap = " Solution of joint approach in Exercise498."}
plotSolution(KETPex498, res)
```

Genotype data is simulated N = 100 times for the missing, transferred to the victims, and the fraction of times the assignment from which the simulation was generated
is found to be 0.93.
```{r, eval = F}
set.seed(1729)
N = 100
missing = KETPex497$missing
am = forrel::profileSim(KETPex497$am, N = N, ids = missing)

res = lapply(am, function(x){
  KETPex497$pm = transferMarkers(x, KETPex497$pm, 
                                 idsFrom = missing, 
                                 idsTo = c("V1", "V2", "V3"))
  jointDVI(KETPex497, verbose = F)
})
correct = lapply(res, function(x)
                all(x[1,1:3] == missing))
correct = unlist(correct)
sum(correct)/N
```
